---
title: "RcppDeepState"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rcppdeepstate-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RcppDeepState)
```

## RcppDeepState

RcppDeepState is an R package that extends the deepstate testing framework to fuzz test the Rcpp packages. DeepState is a static library that provides common interface to fuzzing and symbolic execution engines. Users can create a testharness and execute it under those fuzzing engines. It also provides a built-in fuzzer to generate tests with learned or random inputs.

## Generation of TestHarness

RcppDeepState provides three key R functions to test any Rcpp package. 

**(i) RcppDeepState::deepstate_pkg_create**: This function will generate testharnesses for all the functions in the package with default tests being the functions in the package.

We have implemented RcppDeepState_* random generation function for only 8 data types and we can create the testharness for the package only if the datatypes in the function are from the given list. 

RcppDeepState specific datatypes:

{"RcppDeepState_NumericVector","RcppDeepState_NumericMatrix","RcppDeepState_mat",
"RcppDeepState_string","RcppDeepState_CharacterVector","RcppDeepState_int",
"RcppDeepState_IntegerVector","RcppDeepState_double"}

The RcppDeepState_* random generation functions can be overloaded, Users can specify the range and the size of the input that is to be generated.

**usage:** RcppDeepState::deepstate_pkg_create(package_path)
 
**package_path:** is a character vector with a path to the test package.

The output varies depending upon the package:

* If all the functions in the package have their function's argument datatypes in the above specified datatype list then the creation of tests for that function in the package is successful.The function gives a message saying the number of functions the test harness were created for. The function returns a character vector with the names of the testharness that were created.

* If only few functions in the package have their function's argument datatypes in the specified datatype list then the function prints a message giving the count of testharnesses created and also prints the functions for which we failed to create the testharnesses. The function returns  a character vector with the names of the testharness that were created.

* If none of the functions in the package have their functions datatypes in the specified datatype list then the function returns a names of the all those functions.

Consider the test path to the testSAN package:

```{r}
path <- system.file("testpkgs/testSAN", package = "RcppDeepState")
path
```

The test package has 6 Rcpp functions that are expected to produce valgrind issues. 

## TestHarness Creation For a package

```{r}
list.harness <- RcppDeepState::deepstate_pkg_create(path)
list.harness
```

The function returns the character vector of testharnesses created. This function create a function specific folders in inst/testfiles of the testpackage.

Each function specific folder has the function specific testharness and its corresponding makefiles.

## TestHarness Creation for a Function

To create a TestHarness for a single function in the package all we have to do is make a call to deepstate_fun_create(), with function_name and path to the test package.

```{r}
function_name <- "rcpp_read_out_of_bound"
function.harness <- deepstate_fun_create(path,function_name)
function.harness
```

## TestHarness Compilation & Execution for Package

```{r}
compiled.harness.list <- RcppDeepState::deepstate_harness_compile_run(path,5)
compiled.harness.list
```

This function compiles and execute all the functions in the package. We can specify the duration to run the testharness. Usually time limit should always be greater than zero and default time limits runs the testharness for 5 seconds.

This function returns the list of function's Testharness that are successfully compiled and executed.

## TestHarness Compilation & Execution for Function.

```{r}
fun_name <- "rcpp_read_out_of_bound"
harness.fun <- deepstate_fuzz_fun(path,fun_name)
harness.fun

#Runs the harness specified number of seconds
time.limit.seconds <- 6
harness.timer.fun<- deepstate_fuzz_fun(path,fun_name,time.limit.seconds)
harness.timer.fun
```

This function executes and returns the name of the function if it is successfully executed. We can adjust the time.limit.seconds arguments to vary the duration of executing the testharness.

Compiling and executing the testharness creates 2 folders in the function specific folders. One folder named inputs has the argument values stored in .qs files and the function specific output folder which has the .crash/.fail/.pass files depending on the result they produced when run with the harness. We can use these inputs in the later part to run the testharness under Valgrind.

Inputs files will have the most recent values passed on to the harness. Those .qs files can be read using qread() from qs package.

```{r}
inputs <- Sys.glob(file.path(path,fun_name,"inputs/*"))
#RcppDeepState::list_inputs(inputs)
```

## Analyze Package Using Valgrind

Now that we the inputs that are created by deepstate_pkg_compile() or deepstate_fuzz_fun() we can use these .crash/.fail/.pass as inputs to the testharness when we run it under valgrind.

```{r}
analyzed.harness <- deepstate_harness_analyze_pkg(path)
analyzed.harness
#gives the logtable datatable
RcppDeepState::issues.table(analyzed.harness$logtable)
#gives inputs datatable
RcppDeepState::inputs.table(analyzed.harness$inputs)
```

This function returns a data.table with three columns. The first gives the path to the binary files (.crash/.fail/.pass). The second column gives a list of inputs that are passed to the testharness. The third column has a data.table with list of error messages.

We can also specify the number of binary files to analyze and number of test functions to analyze.

```{r}
analyzed.harness <- deepstate_harness_analyze_pkg(path,2,1)
analyzed.harness
```

This function analyzes two testharness functions with one binary file on each of them.The function's default binary files to analyze is set to 3. And the default number for testfiles is to analyze all.

## Analyze Function Using Valgrind

```{r}
fun_name <- "rcpp_read_out_of_bound"
analyzed.harness.fun <- deepstate_analyze_fun(path,fun_name)
analyzed.harness.fun
```

The function analyzes a function on given number of binary files.The default max_inpus is set to analyze 3 binary files. The number of max_inputs should be less than number of binary files in the function specifi output folder.

We can change the number of max_inputs like this.

```{r}
fun_name <- "rcpp_read_out_of_bound"
analyzed.harness.fun <- deepstate_analyze_fun(path,fun_name)
analyzed.harness.fun
```

## Analyze binary file using Valgrind

We can analyze a single binary file to check the output of the file.

```{r}
fun_name <- "rcpp_read_out_of_bound"
binary.dir <- file.path(package_path,"inst/testfiles",fun_name,paste0(fun_name,"_output"))
binary.files <- Sys.glob(file.path(binary.dir,"*"))
binary.files[1]
analyze.result <- deepstate_analyze_file(binary.files[1])
analyze.result
```
















